Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-07 Manuscript/manuscript.tex
Commit small changes before merge

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
\todo{JOE: list more introductory examples, where it has interesting effects.}
Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
to understand molecular level details of specific cation interactions with
cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} 
\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
While relative binding affinity of different ions is agreed to follow Hoffmeister
series, the molecular details of binding and binding energetics are
not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
%General conclusion from experimental
%studies has been that multivalent ions and lithium have weak specific binding in
%phopholipid bilayers, while other monovalent ions do not essentially
%bind 
%The presence of anoinic lipids, like PS or PG,
%increase the concentration close to the bilayer and thus the amount of bound
%ions, but do not affect the specific binding constant \cite{seelig90}.
Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
and infrared scpectroscopy, give accurate information about ion binding
in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
Molecular level interpretation of the results, however, requires assumptions
about binding models and is often not fully conclusive \cite{??}.
On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
The overestimated specific cation binding observed in current lipid models \cite{catte16}
may lead to articifially positively charged memranes and significant artefacts in
MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
\todo{JOE: this may be a too daring statement. \\
  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
  but it might happen with too small concentrations. 
  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).

>>>>

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
\todo{JOE: list more introductory examples, where it has interesting effects.}

Zwitterionic phosphocholine (PC) lipid bilayers are common model systems for
cellular membranes. Interactions of biological cations, especially Na$^+$ and
Ca$^{2+}$ are with 
have been widely used as
model systems to study details of cation interactions with cellular membranes
by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??}

While relative binding affinity of different ions is agreed to follow Hoffmeister
series, the molecular details of binding and binding energetics are
not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.

%General conclusion from experimental
%studies has been that multivalent ions and lithium have weak specific binding in
%phopholipid bilayers, while other monovalent ions do not essentially
%bind 
%The presence of anoinic lipids, like PS or PG,
%increase the concentration close to the bilayer and thus the amount of bound
%ions, but do not affect the specific binding constant \cite{seelig90}.
Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
and infrared scpectroscopy, give accurate information about ion binding
in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
Molecular level interpretation of the results, however, requires assumptions
about binding models and is often not fully conclusive \cite{??}.
On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
The overestimated specific cation binding observed in current lipid models \cite{catte16}
may lead to articifially positively charged memranes and significant artefacts in
MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
\todo{JOE: this may be a too daring statement. \\
  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
  but it might happen with too small concentrations. 
  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
